---
title: "rs36047 -> miR-584-5p -> exacerbation mediation"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  cra_counts: "../data/CRA/2022-08-16_CRA_1159_exceRpt_miRNA_ReadCountscopy.csv"
  cra_pheno: "../data/CRA/PhenotypeInfo.csv"
  cra_link: "../data/CRA/cra_shortened_phenotype_files10-11.csv"
  cra_snp: "../data/CRA/gacrs_mir584_cis.raw"
  cra_pcs: "../data/CRA/gacrs_pcs.txt"
  camp_counts: "../data/CAMP/CAMP.492.COUNT.csv"
  camp_pheno: "../data/CAMP/camp_pheno_2022-08-04.tsv"
  camp_link: "../data/CAMP/shortened_phenotype_files_ver5.csv"
  camp_snp: "../data/CAMP/camp_mir584_cis.raw"
  camp_pcs: "../data/CAMP/camp_pcs.txt"
  outdir: "../results"
  highlight_mir: "hsa-miR-584-5p"
  min_count: 5
  min_prop_samples: 0.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(stringsAsFactors = FALSE)
```

```{r packages}
library(dplyr)
library(ggplot2)
library(preprocessCore)
library(broom)
library(mediation)
library(stringr)
```

## Output folders

```{r output-folders}
outdir <- params$outdir

med_tables_dir <- file.path(outdir, "mediation", "tables")
med_plots_dir  <- file.path(outdir, "mediation", "plots")

dir.create(med_tables_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(med_plots_dir, recursive = TRUE, showWarnings = FALSE)
```

## Helper: quantile-normalize one miRNA

```{r helper-functions}
extract_miR584_quantile <- function(count_df, mir_col, mir_name = "hsa-miR-584-5p",
                                    min_count = 5, min_prop = 0.5) {
  keep <- rowSums(count_df[, -1] > min_count) >= (ncol(count_df[, -1]) * min_prop)
  miRNAs <- count_df[keep, ]

  mir_mat <- as.matrix(miRNAs[, -1])
  rownames(mir_mat) <- miRNAs[[mir_col]]

  mir_norm <- normalize.quantiles(mir_mat)
  rownames(mir_norm) <- rownames(mir_mat)
  colnames(mir_norm) <- colnames(mir_mat)

  out <- as.data.frame(mir_norm[mir_name, ])
  out$svid <- rownames(out)
  rownames(out) <- NULL
  colnames(out)[1] <- "miRNA584"

  out
}
```

# CAMP

## CAMP input files

```{r camp-input}
camp_snp   <- read.table(params$camp_snp, header = TRUE)
camp_link  <- read.csv(params$camp_link)
camp_counts <- read.csv(params$camp_counts)
camp_pheno <- read.table(params$camp_pheno, sep = "\t", header = TRUE)
camp_pcs   <- read.table(params$camp_pcs, header = TRUE)
```

## CAMP miRNA normalization

```{r camp-mirna}
camp_miR584 <- extract_miR584_quantile(
  count_df   = camp_counts,
  mir_col    = "MIRNA",
  mir_name   = params$highlight_mir,
  min_count  = params$min_count,
  min_prop   = params$min_prop_samples
)
```

## CAMP prepare and merge

```{r camp-merge}
camp_pheno2 <- camp_pheno %>%
  filter(TG != 2) %>%
  mutate(across(c(camp, U_PEDIGREEID, S_SUBJECTID, TG), as.character))

camp_link2 <- camp_link %>%
  mutate(across(c(CAMP_ID, row.names.CAMP_492_COUNT.Trans.), as.character))

camp_miR584 <- camp_miR584 %>%
  mutate(svid = as.character(svid))
colnames(camp_snp) <- trimws(colnames(camp_snp))
camp_snp2 <- camp_snp %>%
  dplyr::rename(pedid = FID, subid = IID) %>%
  mutate(across(c(pedid, subid), as.character))

camp_map <- camp_pheno2 %>%
  dplyr::rename(
    campid = camp,
    pedid = U_PEDIGREEID,
    subid = S_SUBJECTID
  ) %>%
  dplyr::select(campid, pedid, subid, TG, AGE, SEX, EDHOS_cum_Y1)

camp_link_std <- camp_link2 %>%
  dplyr::rename(
    campid = CAMP_ID,
    svid = row.names.CAMP_492_COUNT.Trans.
  )

camp_pcs$subid <- as.character(camp_pcs$S_SUBJECTID)

camp_df <- camp_link_std %>%
  left_join(camp_map, by = "campid") %>%
  inner_join(camp_pcs, by = "subid") %>%
  inner_join(camp_snp2, by = "subid") %>%
  inner_join(camp_miR584, by = "svid")

camp_df$treatment_grp <- ifelse(camp_df$TG == 1, "ICS", "Placebo")
camp_df$treatment_grp <- relevel(factor(camp_df$treatment_grp), ref = "Placebo")

camp_df$miRNA584_norm <- qnorm((rank(camp_df$miRNA584) - 0.5) / length(camp_df$miRNA584))
```

## CAMP define outcome and SNP


```{r camp-outcome}
lead_snp_camp <- c("X5.149061758.C.T_C","X5.149031357.A.C_A", "X5.149121003.G.A_A")

camp_df <- camp_df |>
  filter(!is.na(EDHOS_cum_Y1)) |>
  mutate(
    outcome = factor(ifelse(EDHOS_cum_Y1 < 1, "low", "high")),
    outcome = relevel(outcome, ref = "low"),
    treatment = factor(TG),
    treatment = relevel(treatment, ref = "3"),
    SEX.x = factor(SEX.x)
  )
# stopifnot(lead_snp_camp %in% names(camp_df))
```

## CAMP SNP -> exacerbation check

```{r camp-snp-to-outcome}
results_list <- list()

for (snp in lead_snp_camp) {

  # build formula safely
  formula_str <- paste0(
    "outcome ~ `", snp, "` + AGE.x + SEX.x + ",
    "PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"
  )

  model <- glm(
    as.formula(formula_str),
    data = camp_df,
    family = binomial()
  )

  # extract only SNP row
  res <- broom::tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
    filter(term == snp) %>%
    mutate(SNP = snp)

  results_list[[snp]] <- res
}

# combine all results
final_results <- bind_rows(results_list)

final_results
```

## CAMP SNP -> exacerbation check for ICS subgroup

```{r camp-snp-to-outcome}
results_list <- list()
ics_df = camp_df[camp_df$TG==1,]
for (snp in lead_snp_camp) {

  # build formula safely
  formula_str <- paste0(
    "outcome ~ `", snp, "` + AGE.x + SEX.x + ",
    "PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"
  )

  model <- glm(
    as.formula(formula_str),
    data = ics_df,
    family = binomial()
  )

  # extract only SNP row
  res <- broom::tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
    filter(term == snp) %>%
    mutate(SNP = snp)

  results_list[[snp]] <- res
}

# combine all results
ics_results <- bind_rows(results_list)

ics_results
```


## CAMP SNP -> miRNA mediator model

```{r camp-mediator-model}
results_list <- list()
covars <- c("AGE.x", "SEX.x", paste0("PC", 1:10))

for (snp in lead_snp_camp) {

  # build formula safely
  formula_str <- paste0(
    "miRNA584_norm ~ `", snp, "` + AGE.x + SEX.x + treatment_grp + ",
    "PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"
  )

  camp_med_fit <- lm(
    as.formula(formula_str),
    data = camp_df
  )
  # camp_med_fit <- lm(
  # miRNA584_norm ~ `X5.149061758.C.T_C` + AGE.x + SEX.x + treatment_grp +
  #   PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  # data = camp_df
  # )

  # extract 
  res <- broom::tidy(camp_med_fit, conf.int = TRUE) %>%
    filter(!term %in% covars) %>%   # remove covariates
    mutate(SNP_tested = snp)

  results_list[[snp]] <- res
}

# combine all results
final_results <- bind_rows(results_list)

final_results

```

## CAMP miRNA -> exacerbation outcome model with interaction

```{r camp-outcome-model-interaction}
covars <- c("AGE.x", "SEX.x", paste0("PC", 1:10))

results_list <- list()
model_list <- list()

for (snp in lead_snp_camp) {

  formula_str <- paste0(
    "outcome ~ `", snp, "` + miRNA584_norm * treatment_grp + ",
    "AGE.x + SEX.x + ",
    "PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"
  )

  fit <- glm(
    as.formula(formula_str),
    data = camp_df,
    family = binomial()
  )

  model_list[[snp]] <- fit

  res <- broom::tidy(fit, exponentiate = TRUE, conf.int = TRUE) %>%
    filter(!term %in% covars) %>%   # remove covariates
    mutate(SNP_tested = snp)

  results_list[[snp]] <- res
}

final_results_int <- bind_rows(results_list)

final_results_int
```

## CAMP pooled mediation without interaction


```{r camp-mediation-pooled}
camp_out_fit_pooled <- glm(
  outcome ~ `X5.149061758.C.T_C` + miRNA584_norm +
    AGE.x + SEX.x + treatment_grp +
    PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = camp_df,
  family = binomial()
)

camp_med_pooled <- mediate(
  model.m = camp_med_fit,
  model.y = camp_out_fit_pooled,
  treat = "X5.149061758.C.T_C",
  mediator = "miRNA584_norm",
  boot = TRUE,
  sims = 1000
)

summary(camp_med_pooled)
```

## CAMP stratified mediation: ICS only

```{r camp-ics-only}
camp_df_ics <- camp_df %>% filter(treatment_grp == "ICS")

camp_med_fit_ics <- lm(
  miRNA584_norm ~ `X5.149061758.C.T_C` + AGE.x + SEX.x +
    PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = camp_df_ics
)

camp_out_fit_ics <- glm(
  outcome ~ `X5.149061758.C.T_C` + miRNA584_norm +
    AGE.x + SEX.x +
    PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = camp_df_ics,
  family = binomial()
)

camp_med_ics <- mediate(
  model.m = camp_med_fit_ics,
  model.y = camp_out_fit_ics,
  treat = "X5.149061758.C.T_C",
  mediator = "miRNA584_norm",
  boot = TRUE,
  sims = 1000
)

summary(camp_med_ics)
```

## CAMP stratified mediation: Placebo only

```{r camp-placebo-only}
camp_df_placebo <- camp_df %>% filter(treatment_grp == "Placebo")

camp_med_fit_placebo <- lm(
  miRNA584_norm ~ `X5.149061758.C.T_C` + AGE.x + SEX.x +
    PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = camp_df_placebo
)

camp_out_fit_placebo <- glm(
  outcome ~ `X5.149061758.C.T_C` + miRNA584_norm +
    AGE.x + SEX.x +
    PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = camp_df_placebo,
  family = binomial()
)

camp_med_placebo <- mediate(
  model.m = camp_med_fit_placebo,
  model.y = camp_out_fit_placebo,
  treat = "X5.149061758.C.T_C",
  mediator = "miRNA584_norm",
  boot = TRUE,
  sims = 1000
)

summary(camp_med_placebo)
```

# CRA / GACRS

## CRA input files

```{r cra-input}
cra_snp   <- read.table(params$cra_snp, header = TRUE)
cra_link  <- read.csv(params$cra_link)
cra_counts <- read.csv(params$cra_counts)
cra_pheno <- read.table(params$cra_pheno, sep = ",", header = TRUE)
cra_pcs   <- read.table(params$cra_pcs, header = TRUE)
```

## CRA miRNA normalization

```{r cra-mirna}
cra_miR584 <- extract_miR584_quantile(
  count_df   = cra_counts,
  mir_col    = "X",
  mir_name   = params$highlight_mir,
  min_count  = params$min_count,
  min_prop   = params$min_prop_samples
)
```

## CRA prepare and merge

```{r cra-merge}
cra_pheno2 <- cra_pheno %>%
  mutate(across(c(studyid, S_SUBJECTID, Inhaled_Steroids), as.character))

cra_link2 <- cra_link %>%
  mutate(across(c(ST.ID, sample.id), as.character))

cra_link2$sample.id <- gsub("\\-", ".", cra_link2$sample.id)

cra_miR584 <- cra_miR584 %>%
  mutate(svid = as.character(svid))

cra_snp2 <- cra_snp %>%
  rename(pedid = FID, subid = IID) %>%
  mutate(across(c(pedid, subid), as.character)) %>%
  distinct(subid, .keep_all = TRUE)

cra_map <- cra_pheno2 %>%
  dplyr::rename(
    subid = S_SUBJECTID,
    TG = Inhaled_Steroids
  ) %>%
  dplyr::select(studyid, subid, TG, Hospitalized_Asthma_Last_Yr)

cra_link_std <- cra_link2 %>%
  rename(
    subid = ST.ID,
    svid = sample.id
  ) %>%
  distinct(subid, svid, .keep_all = TRUE)

cra_pcs$subid <- as.character(cra_pcs$IID)

cra_df <- cra_link_std %>%
  left_join(cra_map, by = "subid") %>%
  inner_join(cra_snp2, by = "subid") %>%
  inner_join(cra_pcs, by = "subid") %>%
  inner_join(cra_miR584, by = "svid")

cra_df$treatment_grp <- ifelse(cra_df$TG == 2, "ICS", "Placebo")
cra_df$treatment_grp <- relevel(factor(cra_df$treatment_grp), ref = "Placebo")

cra_df$miRNA584_norm <- qnorm((rank(cra_df$miRNA584) - 0.5) / length(cra_df$miRNA584))
```

## CRA define outcome and SNP

```{r cra-outcome}
lead_snp_cra <- "X5.149061758.C.T_C"

cra_df <- cra_df |>
  filter(!is.na(Hospitalized_Asthma_Last_Yr)) |>
  mutate(
    outcome = factor(Hospitalized_Asthma_Last_Yr),
    treatment = factor(TG),
    treatment = relevel(treatment, ref = "1"),
    gender.x = factor(gender)
  )

stopifnot(lead_snp_cra %in% names(cra_df))
```

## CRA SNP -> exacerbation check

```{r cra-snp-to-outcome}
cra_snp_outcome <- glm(
  outcome ~ `X5.149061758.C.T_C` + age + gender + 
    PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = cra_df,
  family = binomial()
)

summary(cra_snp_outcome)
broom::tidy(cra_snp_outcome, exponentiate = TRUE, conf.int = TRUE)
```

## CRA SNP -> miRNA mediator model

```{r cra-mediator-model}
cra_med_fit <- lm(
  miRNA584_norm ~ `X5.149061758.C.T_C` + age + gender + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10 + treatment_grp,
  data = cra_df
)

summary(cra_med_fit)
broom::tidy(cra_med_fit)
```

## CRA miRNA -> exacerbation outcome model with interaction

```{r cra-outcome-model-interaction}
cra_out_fit_int <- glm(
  outcome ~ `X5.149061758.C.T_C` + miRNA584_norm * treatment_grp + age + gender + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = cra_df,
  family = binomial()
)

summary(cra_out_fit_int)
broom::tidy(cra_out_fit_int, exponentiate = TRUE, conf.int = TRUE)
```

## CRA pooled mediation without interaction

```{r cra-mediation-pooled}
cra_out_fit_pooled <- glm(
  outcome ~ `X5.149061758.C.T_C` + miRNA584_norm + treatment_grp + age + gender + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = cra_df,
  family = binomial()
)

cra_med_pooled <- mediate(
  model.m = cra_med_fit,
  model.y = cra_out_fit_pooled,
  treat = "X5.149061758.C.T_C",
  mediator = "miRNA584_norm",
  boot = TRUE,
  sims = 1000
)

summary(cra_med_pooled)
```

## CRA stratified mediation: ICS only

```{r cra-ics-only}
cra_df_ics <- cra_df %>% filter(treatment_grp == "ICS")

cra_med_fit_ics <- lm(
  miRNA584_norm ~ `X5.149061758.C.T_C` + age + gender + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = cra_df_ics
)

cra_out_fit_ics <- glm(
  outcome ~ `X5.149061758.C.T_C` + miRNA584_norm + age + gender + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = cra_df_ics,
  family = binomial()
)

cra_med_ics <- mediate(
  model.m = cra_med_fit_ics,
  model.y = cra_out_fit_ics,
  treat = "X5.149061758.C.T_C",
  mediator = "miRNA584_norm",
  boot = TRUE,
  sims = 1000
)

summary(cra_med_ics)
```

## CRA stratified mediation: Placebo only

```{r cra-placebo-only}
cra_df_placebo <- cra_df %>% filter(treatment_grp == "Placebo")

cra_med_fit_placebo <- lm(
  miRNA584_norm ~ `X5.149061758.C.T_C` + age + gender + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = cra_df_placebo
)

cra_out_fit_placebo <- glm(
  outcome ~ `X5.149061758.C.T_C` + miRNA584_norm + age + gender + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10,
  data = cra_df_placebo,
  family = binomial()
)

cra_med_placebo <- mediate(
  model.m = cra_med_fit_placebo,
  model.y = cra_out_fit_placebo,
  treat = "X5.149061758.C.T_C",
  mediator = "miRNA584_norm",
  boot = TRUE,
  sims = 1000
)

summary(cra_med_placebo)
```

# Save key outputs

```{r save-key-outputs}
write.csv(broom::tidy(camp_snp_outcome, exponentiate = TRUE, conf.int = TRUE),
          file.path(med_tables_dir, "CAMP_rs36047_snp_to_exacerbation.csv"),
          row.names = FALSE)

write.csv(broom::tidy(camp_med_fit),
          file.path(med_tables_dir, "CAMP_rs36047_mediator_model.csv"),
          row.names = FALSE)

write.csv(broom::tidy(camp_out_fit_int, exponentiate = TRUE, conf.int = TRUE),
          file.path(med_tables_dir, "CAMP_rs36047_outcome_model_interaction.csv"),
          row.names = FALSE)

write.csv(broom::tidy(cra_snp_outcome, exponentiate = TRUE, conf.int = TRUE),
          file.path(med_tables_dir, "CRA_rs36047_snp_to_exacerbation.csv"),
          row.names = FALSE)

write.csv(broom::tidy(cra_med_fit),
          file.path(med_tables_dir, "CRA_rs36047_mediator_model.csv"),
          row.names = FALSE)

write.csv(broom::tidy(cra_out_fit_int, exponentiate = TRUE, conf.int = TRUE),
          file.path(med_tables_dir, "CRA_rs36047_outcome_model_interaction.csv"),
          row.names = FALSE)
```

# Summary objects from mediation output

```{r mediation-summary-objects}
camp_med_ics_sum <- summary(camp_med_ics)
camp_med_placebo_sum <- summary(camp_med_placebo)
cra_med_ics_sum <- summary(cra_med_ics)
cra_med_placebo_sum <- summary(cra_med_placebo)
```



