---
title: "mirna_cis_miR_QTL"
output: html_document
date: "2026-03-13"

params:
  cra_counts: "../data/CRA/2022-08-16_CRA_1159_exceRpt_miRNA_ReadCountscopy.csv"
  cra_pheno: "../data/CRA/PhenotypeInfo.csv"
  cra_link: "../data/CRA/cra_shortened_phenotype_files10-11.csv"
  camp_counts: "../data/CAMP/CAMP.492.COUNT.csv"
  camp_pheno: "../data/CAMP/camp_pheno_2022-08-04.tsv"
  camp_link: "../data/CAMP/shortened_phenotype_files_ver5.csv"
  outdir: "../results"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(stringsAsFactors = FALSE)
# rm(list = ls())
```

## Libraries

```{r libraries}
library(dplyr)
library(broom)
library(preprocessCore)
library(ggplot2)
library(stringr)
```

```{r paths}
outdir <- params$outdir
dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "QTL"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "QTL_followup"), recursive = TRUE, showWarnings = FALSE)
```

## CAMP input files

```{r camp-input-files}
SNP   <- read.table("../data/CAMP/camp_mir584_cis.raw", header = TRUE)
covar <- read.csv("../data/CAMP/shortened_phenotype_files_ver5.csv")
miRs  <- read.csv("../data/CAMP/CAMP.492.COUNT.csv")
mapping <- read.table("../data/CAMP/camp_pheno_2022-08-04.tsv", sep = "\t", header = TRUE)
pcs <- read.table("../data/CAMP/camp_pcs.txt", header = TRUE)
```

## CAMP miRNA normalization

```{r camp-mirna-normalization}
miRNAs <- miRs[rowSums(miRs > 5) >= (ncol(miRs) * 0.5), ]

mir_mat <- as.matrix(miRNAs[, -1])
rownames(mir_mat) <- miRNAs$MIRNA

miR_norm <- normalize.quantiles(mir_mat)
rownames(miR_norm) <- rownames(mir_mat)
colnames(miR_norm) <- colnames(mir_mat)

miR584 <- as.data.frame(miR_norm["hsa-miR-584-5p", ])
miR584$svid <- rownames(miR584)
rownames(miR584) <- NULL
colnames(miR584)[1] <- "miRNA584"
```

## CAMP prepare phenotype and covariate data

```{r camp-prepare-phenotype}
mapping <- mapping[mapping$TG != 2, ]
mapping <- mapping %>% mutate(across(c(camp, U_PEDIGREEID, S_SUBJECTID, TG), as.character))
covar <- covar %>% mutate(across(c(CAMP_ID, row.names.CAMP_492_COUNT.Trans.), as.character))
miR584 <- miR584 %>% mutate(svid = as.character(svid))
```

## CAMP standardize IDs

```{r camp-standardize-ids}
mapping_std <- mapping %>%
  rename(
    campid = camp,
    pedid  = U_PEDIGREEID,
    subid  = S_SUBJECTID,
    TG     = TG
  ) %>%
  dplyr::select(campid, pedid, subid, TG, AGE, SEX)

covar_std <- covar %>%
  rename(
    campid = CAMP_ID,
    svid   = row.names.CAMP_492_COUNT.Trans.
  )

snp2 <- SNP %>%
  rename(pedid = FID, subid = IID) %>%
  mutate(across(c(pedid, subid), as.character))
```

## CAMP merge data

```{r camp-merge-data}
covar_map <- covar_std %>%
  left_join(mapping_std, by = "campid")

pcs$subid <- pcs$S_SUBJECTID

df <- covar_map %>%
  inner_join(pcs, by = "subid") %>%
  inner_join(snp2, by = "subid") %>%
  inner_join(miR584, by = "svid") 

df$treatment_grp <- ifelse(df$TG == 1, "ICS", "Placebo")
df$treatment_grp <- relevel(factor(df$treatment_grp), ref = "Placebo")
df$miRNA584_norm <- qnorm((rank(df$miRNA584) - 0.5) / length(df$miRNA584))
```

## CAMP quick checks

```{r camp-quick-checks}
dim(df)
table(df$treatment_grp)
summary(df$miRNA584_norm)
```

## CAMP identify SNP columns

```{r camp-snp-columns}
snp_cols <- setdiff(
  names(snp2),
  c("pedid", "subid", "PAT", "MAT", "SEX", "PHENOTYPE")
)
length(snp_cols)
head(snp_cols)
```


## CAMP base model: miRNA584 ~ SNP + covariates

```{r camp-base-model-scan}
results_base <- data.frame()

for (snp in snp_cols) {

  snp_bt <- paste0("`", snp, "`")

  model_formula <- as.formula(
    paste0(
      "miRNA584_norm ~ ", snp_bt,
      " + AGE.x + SEX.x",
      " + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"
    )
  )

  fit <- try(lm(model_formula, data = df), silent = TRUE)
  if (inherits(fit, "try-error")) next

  tmp <- broom::tidy(fit)

  snp_row <- tmp[tmp$term == snp, ]
  if (nrow(snp_row) == 0) next

  results_base <- rbind(
    results_base,
    data.frame(
      SNP = snp,
      beta_snp = snp_row$estimate,
      p_snp = snp_row$p.value
    )
  )
}

if (nrow(results_base) > 0) {
  results_base$p_snp_fdr <- p.adjust(results_base$p_snp, method = "fdr")
  results_base <- results_base %>% arrange(p_snp)
}

head(results_base, 20)
```

## CAMP interaction model: miRNA584 ~ SNP + ICS + SNP*ICS + covariates

```{r camp-interaction-model-scan}
results_int <- data.frame()

for (snp in snp_cols) {

  snp_bt <- paste0("`", snp, "`")

  model_formula <- as.formula(
    paste0(
      "miRNA584_norm ~ ", snp_bt,
      " + treatment_grp + ",
      snp, ":treatment_grp",
      " + AGE.x + SEX.x",
      " + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"
    )
  )

  fit <- try(lm(model_formula, data = df), silent = TRUE)
  if (inherits(fit, "try-error")) next

  tmp <- broom::tidy(fit)

  snp_row <- tmp[tmp$term == snp, ]
  int_row <- tmp[tmp$term == paste0(snp, ":treatment_grpICS"), ]

  if (nrow(snp_row) == 0) next

  if (nrow(int_row) == 0) {
    int_beta <- NA
    int_p <- NA
  } else {
    int_beta <- int_row$estimate
    int_p <- int_row$p.value
  }

  results_int <- rbind(
    results_int,
    data.frame(
      SNP = snp,
      beta_snp = snp_row$estimate,
      p_snp = snp_row$p.value,
      beta_int = int_beta,
      p_int = int_p
    )
  )
}

results_int$p_snp_fdr <- p.adjust(results_int$p_snp, method = "fdr")
results_int$p_int_fdr <- p.adjust(results_int$p_int, method = "fdr")

results_int <- results_int %>% arrange(p_int, p_snp)

head(results_int, 20)
```

## CAMP simple regional plot

```{r camp-regional-plot}
camp_plot_df <- results_int
camp_plot_df$index <- 1:nrow(camp_plot_df)
camp_plot_df$logp_int <- -log10(camp_plot_df$p_int)

ggplot(camp_plot_df, aes(x = index, y = logp_int)) +
  geom_point(alpha = 0.7) +
  labs(
    title = "CAMP: MIR584 cis-mirQTL interaction scan",
    x = "SNP index across extracted region",
    y = "-log10(p interaction)"
  ) +
  theme_bw()
```
```{r}
results_base$pos <- as.numeric(sub("X5\\.(\\d+)\\..*", "\\1", results_base$SNP))

ggplot(results_base, aes(pos, -log10(p_snp))) +
  geom_point() +
  theme_bw() +
  labs(x="chr5 position", y="-log10(p)")
```

## CAMP save results

```{r camp-save-results}
write.csv(results_base,
          "../results/QTL/MIR584_cis_mirQTL_base_model_CAMP.csv",
          row.names = FALSE)

write.csv(results_int,
          "../results/QTL/MIR584_cis_mirQTL_interaction_model_CAMP.csv",
          row.names = FALSE)
```

## CRA input files

```{r cra-input-files}
SNP_cra   <- read.table("../data/CRA/gacrs_mir584_cis.raw", header = TRUE)
covar_cra <- read.csv("../data/CRA/cra_shortened_phenotype_files10-11.csv")
miRs_cra  <- read.csv("../data/CRA/2022-08-16_CRA_1159_exceRpt_miRNA_ReadCountscopy.csv", sep = ",", header = TRUE)
mapping_cra <- read.table("../data/CRA/PhenotypeInfo.csv", sep = ",", header = TRUE)
```

## CRA miRNA normalization

```{r cra-mirna-normalization}
miRNAs_cra <- miRs_cra[rowSums(miRs_cra > 5) >= (ncol(miRs_cra) * 0.5), ]

mir_mat_cra <- as.matrix(miRNAs_cra[, -1])
rownames(mir_mat_cra) <- miRNAs_cra$X

miR_norm_cra <- normalize.quantiles(mir_mat_cra)
rownames(miR_norm_cra) <- rownames(mir_mat_cra)
colnames(miR_norm_cra) <- colnames(mir_mat_cra)

miR584_cra <- as.data.frame(miR_norm_cra["hsa-miR-584-5p", ])
miR584_cra$svid <- rownames(miR584_cra)
rownames(miR584_cra) <- NULL
colnames(miR584_cra)[1] <- "miRNA584"
```

## CRA prepare phenotype and covariate data

```{r cra-prepare-phenotype}
mapping_cra <- mapping_cra %>%
  mutate(across(c(studyid, S_SUBJECTID, Inhaled_Steroids), as.character))

covar_cra <- covar_cra %>%
  mutate(across(c(ST.ID, sample.id), as.character))

covar_cra$sample.id <- gsub("\\-", ".", covar_cra$sample.id)

miR584_cra <- miR584_cra %>%
  mutate(svid = as.character(svid))
```

## CRA standardize IDs

```{r cra-standardize-ids}
mapping_std_cra <- mapping_cra %>%
  rename(
    subid = S_SUBJECTID,
    TG    = Inhaled_Steroids
  ) %>%
  dplyr::select(studyid, subid, TG)

covar_std_cra <- covar_cra %>%
  rename(
    subid = ST.ID,
    svid  = sample.id
  )

snp2_cra <- SNP_cra %>%
  rename(pedid = FID, subid = IID) %>%
  mutate(across(c(pedid, subid), as.character)) %>%
  distinct(subid, .keep_all = TRUE)

mapping_std_cra <- mapping_std_cra %>% distinct(subid, .keep_all = TRUE)
covar_std_cra   <- covar_std_cra %>% distinct(subid, svid, .keep_all = TRUE)
miR584_cra      <- miR584_cra %>% distinct(svid, .keep_all = TRUE)
```

## CRA merge data

```{r cra-merge-data}
covar_map_cra <- covar_std_cra %>%
  left_join(mapping_std_cra, by = "subid")

df_cra <- covar_map_cra %>%
  inner_join(snp2_cra, by = "subid") %>%
  inner_join(miR584_cra, by = "svid")

df_cra$treatment_grp <- ifelse(df_cra$TG == 2, "ICS", "Placebo")
df_cra$treatment_grp <- relevel(factor(df_cra$treatment_grp), ref = "Placebo")
df_cra$miRNA584_norm <- qnorm((rank(df_cra$miRNA584) - 0.5) / length(df_cra$miRNA584))
```

## CRA quick checks

```{r cra-quick-checks}
dim(df_cra)
table(df_cra$treatment_grp)
summary(df_cra$miRNA584_norm)
```

## CRA identify SNP columns

```{r cra-snp-columns}
snp_cols_cra <- setdiff(names(snp2_cra),
  c("pedid", "subid", "PAT", "MAT", "SEX", "PHENOTYPE"))
length(snp_cols_cra)
head(snp_cols_cra)
```

## CRA base model: miRNA584 ~ SNP + covariates

```{r cra-base-model-scan}
results_base_cra <- data.frame()

for (snp in snp_cols_cra) {

  model_formula <- as.formula(
    paste0(
      "miRNA584_norm ~ ", snp,
      " + age + gender"
    )
  )

  fit <- try(lm(model_formula, data = df_cra), silent = TRUE)
  if (inherits(fit, "try-error")) next

  tmp <- broom::tidy(fit)
  snp_row <- tmp[tmp$term == snp, ]
  if (nrow(snp_row) == 0) next

  results_base_cra <- rbind(
    results_base_cra,
    data.frame(
      SNP = snp,
      beta_snp = snp_row$estimate,
      p_snp = snp_row$p.value
    )
  )
}

results_base_cra$p_snp_fdr <- p.adjust(results_base_cra$p_snp, method = "fdr")
results_base_cra <- results_base_cra %>% arrange(p_snp)

head(results_base_cra, 20)
```

## CRA interaction model: miRNA584 ~ SNP + ICS + SNP*ICS + covariates

```{r cra-interaction-model-scan}
results_int_cra <- data.frame()

for (snp in snp_cols_cra) {

  model_formula <- as.formula(
    paste0(
      "miRNA584_norm ~ ", snp,
      " + treatment_grp + ",
      snp, ":treatment_grp",
      " + age + gender"
    )
  )

  fit <- try(lm(model_formula, data = df_cra), silent = TRUE)
  if (inherits(fit, "try-error")) next

  tmp <- broom::tidy(fit)

  snp_row <- tmp[tmp$term == snp, ]
  int_row <- tmp[tmp$term == paste0(snp, ":treatment_grpICS"), ]

  if (nrow(snp_row) == 0) next

  if (nrow(int_row) == 0) {
    int_beta <- NA
    int_p <- NA
  } else {
    int_beta <- int_row$estimate
    int_p <- int_row$p.value
  }

  results_int_cra <- rbind(
    results_int_cra,
    data.frame(
      SNP = snp,
      beta_snp = snp_row$estimate,
      p_snp = snp_row$p.value,
      beta_int = int_beta,
      p_int = int_p
    )
  )
}

results_int_cra$p_snp_fdr <- p.adjust(results_int_cra$p_snp, method = "fdr")
results_int_cra$p_int_fdr <- p.adjust(results_int_cra$p_int, method = "fdr")

results_int_cra <- results_int_cra %>% arrange(p_int, p_snp)

head(results_int_cra, 20)
```

## CRA simple regional plot

```{r cra-regional-plot}
cra_plot_df <- results_int_cra
cra_plot_df$index <- 1:nrow(cra_plot_df)
cra_plot_df$logp_int <- -log10(cra_plot_df$p_int)

ggplot(cra_plot_df, aes(x = index, y = logp_int)) +
  geom_point(alpha = 0.7) +
  labs(
    title = "CRA: MIR584 cis-mirQTL interaction scan",
    x = "SNP index across extracted region",
    y = "-log10(p interaction)"
  ) +
  theme_bw()
```

```{r}
results_base_cra$pos <- as.numeric(sub("X5\\.(\\d+)\\..*", "\\1", results_base_cra$SNP))

ggplot(results_base_cra, aes(pos, -log10(p_snp))) +
  geom_point() +
  theme_bw() +
  labs(x="chr5 position", y="-log10(p)")
```
## CRA save results

```{r cra-save-results}
write.csv(results_base_cra,
          "../results/QTL/MIR584_cis_mirQTL_base_model_CRA.csv",
          row.names = FALSE)

write.csv(results_int_cra,
          "../results/QTL/MIR584_cis_mirQTL_interaction_model_CRA.csv",
          row.names = FALSE)
```

## Notes

```{r notes}
cat("CAMP base model:\n")
cat("miRNA584 ~ SNP + AGE + SEX + PCs\n\n")

cat("CAMP interaction model:\n")
cat("miRNA584 ~ SNP + treatment_grp + SNP*treatment_grp + AGE + SEX + PCs\n\n")

cat("CRA base model:\n")
cat("miRNA584 ~ SNP + age + gender\n\n")

cat("CRA interaction model:\n")
cat("miRNA584 ~ SNP + treatment_grp + SNP*treatment_grp + age + gender\n\n")

cat("Interpretation:\n")
cat("- base model tests overall SNP association with miR-584-5p\n")
cat("- interaction model tests whether SNP effect differs by ICS status\n")
cat("- focus on p_snp_fdr in base model and p_int_fdr in interaction model\n")
cat("- simple plots show signal pattern across the extracted regional SNP set\n")
```

```{r}
top_camp <- results_base %>% slice_min(p_snp, n = 50)
top_cra  <- results_base_cra %>% slice_min(p_snp, n = 50)

intersect(top_camp$SNP, top_cra$SNP)
```
```{r}
top_camp_pos <- results_base %>%
  slice_min(p_snp, n = 50) %>%
  mutate(
    chr = str_extract(SNP, "(?<=X)\\d+(?=\\.)"),
    pos = str_extract(SNP, "(?<=X\\d\\.)\\d+")
  ) %>%
  mutate(
    chr = as.integer(chr),
    pos = as.integer(pos)
  )

top_cra_pos <- results_base_cra %>%
  slice_min(p_snp, n = 50) %>%
  mutate(
    chr = str_extract(SNP, "(?<=X)\\d+(?=\\.)"),
    pos = str_extract(SNP, "(?<=X\\d\\.)\\d+")
  ) %>%
  mutate(
    chr = as.integer(chr),
    pos = as.integer(pos)
  )

overlap_snps = inner_join(
  top_camp_pos %>% select(SNP_camp = SNP, chr, pos, p_camp = p_snp),
  top_cra_pos  %>% select(SNP_cra  = SNP, chr, pos, p_cra  = p_snp),
  by = c("chr", "pos")
)
```


```{r}
camp_hits <- results_base %>%
  filter(pos %in% overlap_snps$pos) %>%
  select(SNP, pos, beta_snp, p_snp)

cra_hits <- results_base_cra %>%
  filter(pos %in% overlap_snps$pos) %>%
  select(SNP, pos, beta_snp, p_snp)

inner_join(
  camp_hits,
  cra_hits,
  by="pos",
  suffix=c("_camp","_cra")
)
```
```{r}
lead_pos <- c(149061758, 149031357, 149121003)
camp_leads <- results_base %>%
  filter(pos %in% lead_pos) %>%
  arrange(match(pos, lead_pos))

cra_leads <- results_base_cra %>%
  filter(pos %in% lead_pos) %>%
  arrange(match(pos, lead_pos))

camp_leads
cra_leads
```

```{r}
results_int <- results_int %>%
  mutate(
    chr = as.integer(str_extract(SNP, "(?<=X)\\d+(?=\\.)")),
    pos = as.integer(str_extract(SNP, "(?<=X\\d\\.)\\d+"))
  )

results_int_cra <- results_int_cra %>%
  mutate(
    chr = as.integer(str_extract(SNP, "(?<=X)\\d+(?=\\.)")),
    pos = as.integer(str_extract(SNP, "(?<=X\\d\\.)\\d+"))
  )

camp_int_leads <- results_int %>%
  filter(pos %in% lead_pos) %>%
  arrange(match(pos, lead_pos))

cra_int_leads <- results_int_cra %>%
  filter(pos %in% lead_pos) %>%
  arrange(match(pos, lead_pos))

camp_int_leads
cra_int_leads
```

```{r CAMP_plots_miRvsgenotype}

for (snp in camp_leads$SNP) {

  p <- ggplot(df, aes_string(x = paste0("factor(`", snp, "`)"), y = "miRNA584_norm")) +
    geom_boxplot() +
    theme_bw() +
    labs(
      title = paste("CAMP:", snp),
      x = "Genotype",
      y = "miR-584-5p expression"
    )

  print(p)
}

```
```{r CAMP_plots_miRvsgenotype_ICS}
for (snp in camp_leads$SNP) {

  p <- ggplot(df, aes_string(x = paste0("factor(`", snp, "`)"), y = "miRNA584_norm", color = "treatment_grp")) +
    geom_boxplot() +
    theme_bw() +
    labs(
      title = paste("CAMP:", snp, "by ICS"),
      x = "Genotype",
      y = "miR-584-5p expression"
    )

  print(p)
}
```

```{r CRA_plots_miRvsgenotype}
for (snp in cra_leads$SNP) {

  p <- ggplot(df_cra, aes_string(x = paste0("factor(`", snp, "`)"), y = "miRNA584_norm")) +
    geom_boxplot() +
    theme_bw() +
    labs(
      title = paste("CRA:", snp),
      x = "Genotype",
      y = "miR-584-5p expression"
    )

  print(p)
}
```
```{r CRA_plots_miRvsgenotype_ICS}
for (snp in cra_leads$SNP) {

  p <- ggplot(df_cra, aes_string(x = paste0("factor(`", snp, "`)"), y = "miRNA584_norm", color = "treatment_grp")) +
    geom_boxplot() +
    theme_bw() +
    labs(
      title = paste("CRA:", snp, "by ICS"),
      x = "Genotype",
      y = "miR-584-5p expression"
    )

  print(p)
}
```
```{r CAMP_paiwise}
camp_snps <- camp_leads$SNP

cond_results_camp <- data.frame()

pairs_camp <- combn(camp_snps, 2, simplify = FALSE)

for (pair in pairs_camp) {

  snp1 <- pair[1]
  snp2 <- pair[2]

  snp1_bt <- paste0("`", snp1, "`")
  snp2_bt <- paste0("`", snp2, "`")

  model_formula <- as.formula(
    paste0(
      "miRNA584_norm ~ ", snp1_bt, " + ", snp2_bt,
      " + AGE.x + SEX.x",
      " + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"
    )
  )

  fit <- lm(model_formula, data = df)
  tmp <- broom::tidy(fit)

  row1 <- tmp[tmp$term == snp1, ]
  row2 <- tmp[tmp$term == snp2, ]

  cond_results_camp <- rbind(
    cond_results_camp,
    data.frame(
      snp1 = snp1,
      snp2 = snp2,
      beta_snp1 = row1$estimate,
      p_snp1 = row1$p.value,
      beta_snp2 = row2$estimate,
      p_snp2 = row2$p.value
    )
  )
}

cond_results_camp
```

```{r three_snps}
s1 <- paste0("`", camp_snps[1], "`")
s2 <- paste0("`", camp_snps[2], "`")
s3 <- paste0("`", camp_snps[3], "`")

fit3_camp <- lm(
  as.formula(
    paste0(
      "miRNA584_norm ~ ", s1, " + ", s2, " + ", s3,
      " + AGE.x + SEX.x",
      " + PC1 + PC2 + PC3 + PC4 + PC5 + PC6 + PC7 + PC8 + PC9 + PC10"
    )
  ),
  data = df
)

broom::tidy(fit3_camp)
```


```{r CRA_pairwise}
cra_snps <- cra_leads$SNP

cond_results_cra <- data.frame()

pairs_cra <- combn(cra_snps, 2, simplify = FALSE)

for (pair in pairs_cra) {

  snp1 <- pair[1]
  snp2 <- pair[2]

  snp1_bt <- paste0("`", snp1, "`")
  snp2_bt <- paste0("`", snp2, "`")

  model_formula <- as.formula(
    paste0(
      "miRNA584_norm ~ ", snp1_bt, " + ", snp2_bt,
      " + age + gender"
    )
  )

  fit <- lm(model_formula, data = df_cra)
  tmp <- broom::tidy(fit)

  row1 <- tmp[tmp$term == snp1, ]
  row2 <- tmp[tmp$term == snp2, ]

  cond_results_cra <- rbind(
    cond_results_cra,
    data.frame(
      snp1 = snp1,
      snp2 = snp2,
      beta_snp1 = row1$estimate,
      p_snp1 = row1$p.value,
      beta_snp2 = row2$estimate,
      p_snp2 = row2$p.value
    )
  )
}

cond_results_cra
```

```{r}
s1 <- paste0("`", cra_snps[1], "`")
s2 <- paste0("`", cra_snps[2], "`")
s3 <- paste0("`", cra_snps[3], "`")

fit3_cra <- lm(
  as.formula(
    paste0(
      "miRNA584_norm ~ ", s1, " + ", s2, " + ", s3,
      " + age + gender"
    )
  ),
  data = df_cra
)

broom::tidy(fit3_cra)
```

```{r meta-analysis}
# Merge cohorts by genomic position
replicated_hits <- inner_join(
  camp_int_leads %>%
    select(SNP_camp = SNP,chr, pos, beta_int_camp = beta_snp, p_int_camp = p_snp),

  cra_int_leads %>%
    select(SNP_cra = SNP,chr, pos, beta_int_cra = beta_snp, p_int_cra = p_snp),

  by = c("chr", "pos")
)

# Keep same direction effects
replicated_hits <- replicated_hits %>%
  filter(sign(beta_int_camp) == sign(beta_int_cra))

# Sort by best CAMP signal
replicated_hits <- replicated_hits %>%
  arrange(p_int_camp)

replicated_hits <- replicated_hits %>%
  filter(p_int_camp < 0.05 | p_int_cra < 0.05)

meta_input <- replicated_hits %>%
  mutate(
    z_camp = qnorm(p_int_camp / 2, lower.tail = FALSE),
    z_cra  = qnorm(p_int_cra  / 2, lower.tail = FALSE),

    se_camp = abs(beta_int_camp) / z_camp,
    se_cra  = abs(beta_int_cra)  / z_cra,

    var_camp = se_camp^2,
    var_cra  = se_cra^2,

    w_camp = 1 / var_camp,
    w_cra  = 1 / var_cra,

    beta_meta = (beta_int_camp * w_camp + beta_int_cra * w_cra) / (w_camp + w_cra),
    se_meta   = sqrt(1 / (w_camp + w_cra)),

    z_meta = beta_meta / se_meta,
    p_meta = 2 * pnorm(abs(z_meta), lower.tail = FALSE),

    q_stat = w_camp * (beta_int_camp - beta_meta)^2 +
             w_cra  * (beta_int_cra  - beta_meta)^2,

    q_df = 1,
    p_het = pchisq(q_stat, df = q_df, lower.tail = FALSE),

    i2 = pmax(0, (q_stat - q_df) / q_stat) * 100
  ) %>%
  arrange(p_meta)

meta_input
```


```{r save-tables-1}
write.csv(top_camp,
          file.path(out_dir, "top50_CAMP_base.csv"),
          row.names = FALSE)

write.csv(top_cra,
          file.path(out_dir, "top50_CRA_base.csv"),
          row.names = FALSE)

write.csv(top_camp_pos,
          file.path(out_dir, "top50_CAMP_base_with_pos.csv"),
          row.names = FALSE)

write.csv(top_cra_pos,
          file.path(out_dir, "top50_CRA_base_with_pos.csv"),
          row.names = FALSE)

write.csv(overlap_snps,
          file.path(out_dir, "overlap_top50_CAMP_CRA_by_chr_pos.csv"),
          row.names = FALSE)

write.csv(camp_hits,
          file.path(out_dir, "overlap_hits_CAMP.csv"),
          row.names = FALSE)

write.csv(cra_hits,
          file.path(out_dir, "overlap_hits_CRA.csv"),
          row.names = FALSE)

replicated_hits <- inner_join(
  camp_hits,
  cra_hits,
  by = "pos",
  suffix = c("_camp", "_cra")
)

write.csv(replicated_hits,
          file.path(out_dir, "replicated_hits_CAMP_CRA.csv"),
          row.names = FALSE)

write.csv(camp_leads,
          file.path(out_dir, "lead3_CAMP_base.csv"),
          row.names = FALSE)

write.csv(cra_leads,
          file.path(out_dir, "lead3_CRA_base.csv"),
          row.names = FALSE)

write.csv(camp_int_leads,
          file.path(out_dir, "lead3_CAMP_interaction.csv"),
          row.names = FALSE)

write.csv(cra_int_leads,
          file.path(out_dir, "lead3_CRA_interaction.csv"),
          row.names = FALSE)

```

