---
title: "miRNA × ICS interaction analysis"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
params:
  cra_counts: "../data/CRA/2022-08-16_CRA_1159_exceRpt_miRNA_ReadCountscopy.csv"
  cra_pheno: "../data/CRA/PhenotypeInfo.csv"
  cra_link: "../data/CRA/cra_shortened_phenotype_files10-11.csv"
  camp_counts: "../data/CAMP/CAMP.492.COUNT.csv"
  camp_pheno: "../data/CAMP/camp_pheno_2022-08-04.tsv"
  camp_link: "../data/CAMP/shortened_phenotype_files_ver5.csv"
  outdir: "../results"
  highlight_mir: "hsa-miR-584-5p"
  min_count: 5
  min_prop_samples: 0.5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
set.seed(2025)
# rm(list=ls())
```

# Overview

This workflow compares normalization methods for circulating miRNA count data,
generates cohort-level demographic summaries, and performs miRNA × ICS interaction
analysis with matched stratified models in CRA and CAMP.

```{r packages}
library(edgeR)
library(preprocessCore)
library(DESeq2)
library(limma)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggrepel)
library(pROC)
library(tibble)
library(purrr)
library(readr)
library(stringr)
library(forcats)
library(patchwork)
library(matrixStats)
library(scales)
```

```{r paths}
outdir <- params$outdir
dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "qc"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "demographics"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "models"), recursive = TRUE, showWarnings = FALSE)
dir.create(file.path(outdir, "figures"), recursive = TRUE, showWarnings = FALSE)
```

# Helper functions

```{r helpers}
filter_mirnas <- function(count_df, min_count = 5, min_prop_samples = 0.5) {
  mat <- as.matrix(count_df[, -1, drop = FALSE])
  rownames(mat) <- count_df[[1]]
  class(mat) <- "numeric"
  keep <- rowSums(mat > min_count) >= (ncol(mat) * min_prop_samples)
  mat[keep, , drop = FALSE]
}

norm_raw_log <- function(mat) log2(mat + 1)

norm_tmm_log <- function(mat) {
  dge <- DGEList(counts = mat)
  dge <- calcNormFactors(dge, method = "TMM")
  log2(cpm(dge, log = FALSE, prior.count = 0.5) + 1)
}

norm_deseq2_log <- function(mat) {
  coldata <- data.frame(row.names = colnames(mat), group = factor(rep("all", ncol(mat))))
  dds <- DESeqDataSetFromMatrix(countData = round(mat), colData = coldata, design = ~1)
  dds <- estimateSizeFactors(dds)
  normalized <- counts(dds, normalized = TRUE)
  log2(normalized + 1)
}

norm_quantile_log <- function(mat) {
  qn <- normalize.quantiles(as.matrix(mat))
  rownames(qn) <- rownames(mat)
  colnames(qn) <- colnames(mat)
  log2(qn + 1)
}

get_rle_df <- function(log_mat, method_name, cohort_name) {
  med <- rowMedians(log_mat, na.rm = TRUE)
  rle_mat <- sweep(log_mat, 1, med, FUN = "-")
  as.data.frame(rle_mat) |>
    rownames_to_column("miRNA") |>
    pivot_longer(-miRNA, names_to = "sample_id", values_to = "rle") |>
    mutate(method = method_name, cohort = cohort_name)
}

get_mean_sd_df <- function(log_mat, method_name, cohort_name) {
  tibble(
    miRNA = rownames(log_mat),
    mean_expr = rowMeans(log_mat, na.rm = TRUE),
    sd_expr = matrixStats::rowSds(log_mat, na.rm = TRUE),
    method = method_name,
    cohort = cohort_name
  )
}

get_density_df <- function(log_mat, method_name, cohort_name) {
  as.data.frame(log_mat) |>
    rownames_to_column("miRNA") |>
    pivot_longer(-miRNA, names_to = "sample_id", values_to = "expression") |>
    mutate(method = method_name, cohort = cohort_name)
}

run_pca_df <- function(log_mat, method_name, cohort_name, pheno_df, treatment_col) {
  pca <- prcomp(t(log_mat), center = TRUE, scale. = TRUE)
  out <- as.data.frame(pca$x[, 1:2, drop = FALSE]) |>
    rownames_to_column("sample_id") |>
    left_join(pheno_df |> rownames_to_column("sample_id"), by = "sample_id") |>
    mutate(method = method_name, cohort = cohort_name, treatment = .data[[treatment_col]])
  attr(out, "variance_explained") <- summary(pca)$importance[2, 1:2]
  out
}

plot_density <- function(df, title_text) {
  ggplot(df, aes(x = expression, group = sample_id, color = sample_id)) +
    geom_density(alpha = 0.35) +
    facet_wrap(~ method, scales = "free") +
    labs(title = title_text, x = "log2 expression", y = "Density") +
    theme_bw(base_size = 12) +
    theme(legend.position = "none")
}

plot_rle <- function(df, title_text) {
  ggplot(df, aes(x = sample_id, y = rle)) +
    geom_boxplot(outlier.size = 0.15) +
    facet_wrap(~ method, scales = "free_x") +
    labs(title = title_text, x = "Sample", y = "Relative log expression") +
    theme_bw(base_size = 12) +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
}

plot_mean_sd <- function(df, title_text) {
  ggplot(df, aes(x = mean_expr, y = sd_expr)) +
    geom_point(alpha = 0.45, size = 0.8) +
    geom_smooth(se = FALSE, linewidth = 0.8) +
    facet_wrap(~ method, scales = "free") +
    labs(title = title_text, x = "Mean log2 expression", y = "SD") +
    theme_bw(base_size = 12)
}

plot_pca <- function(df, title_text) {
  ggplot(df, aes(x = PC1, y = PC2, color = treatment)) +
    geom_point(size = 2, alpha = 0.85) +
    facet_wrap(~ method, scales = "free") +
    labs(title = title_text, x = "PC1", y = "PC2", color = "Treatment") +
    theme_bw(base_size = 12)
}

make_corr_plot <- function(mat1, mat2, label1, label2, cohort_name) {
  common_rows <- intersect(rownames(mat1), rownames(mat2))
  common_cols <- intersect(colnames(mat1), colnames(mat2))
  x <- as.vector(mat1[common_rows, common_cols, drop = FALSE])
  y <- as.vector(mat2[common_rows, common_cols, drop = FALSE])
  corr <- cor(x, y, use = "complete.obs", method = "pearson")

  df <- tibble(x = x, y = y)
  ggplot(df, aes(x = x, y = y)) +
    geom_point(alpha = 0.12, size = 0.4) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
    annotate("text", x = quantile(x, 0.1), y = quantile(y, 0.95),
             label = paste0("r = ", round(corr, 3)), hjust = 0, size = 4) +
    labs(
      title = paste0(cohort_name, ": ", label1, " vs ", label2),
      x = label1,
      y = label2
    ) +
    theme_bw(base_size = 12)
}

summarize_continuous_by_group <- function(df, vars, group_col) {
  map_dfr(vars, function(v) {
    sub <- df |> select(all_of(group_col), all_of(v)) |> filter(!is.na(.data[[v]]))
    pval <- tryCatch(t.test(sub[[v]] ~ sub[[group_col]])$p.value, error = function(e) NA_real_)
    sub |>
      group_by(.data[[group_col]]) |>
      summarise(n = n(), mean = mean(.data[[v]], na.rm = TRUE), sd = sd(.data[[v]], na.rm = TRUE), .groups = "drop") |>
      mutate(variable = v, p_value = pval) |>
      relocate(variable)
  })
}

summarize_categorical_by_group <- function(df, vars, group_col) {
  map_dfr(vars, function(v) {
    tab <- table(df[[group_col]], df[[v]], useNA = "no")
    pval <- tryCatch(chisq.test(tab)$p.value, error = function(e) NA_real_)
    as.data.frame(tab) |>
      rename(group = Var1, level = Var2, n = Freq) |>
      mutate(variable = v, p_value = pval) |>
      relocate(variable)
  })
}

plot_demo_violin <- function(df, vars, labels, group_col, title_text) {
  long_df <- df |>
    select(all_of(group_col), all_of(vars)) |>
    pivot_longer(cols = all_of(vars), names_to = "variable", values_to = "value")

  ggplot(long_df, aes(x = .data[[group_col]], y = value, fill = .data[[group_col]])) +
    geom_violin(trim = FALSE, alpha = 0.4, color = NA) +
    geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.7) +
    facet_wrap(~ variable, scales = "free_y", labeller = labeller(variable = labels)) +
    labs(title = title_text, x = NULL, y = "Value") +
    theme_bw(base_size = 12) +
    theme(legend.position = "none")
}

extract_glm_term <- function(model, term_name) {
  coef_tab <- summary(model)$coefficients
  if (!term_name %in% rownames(coef_tab)) {
    return(tibble(term = term_name, estimate = NA_real_, se = NA_real_, p_value = NA_real_, or = NA_real_))
  }
  tibble(
    term = term_name,
    estimate = coef_tab[term_name, "Estimate"],
    se = coef_tab[term_name, "Std. Error"],
    p_value = coef_tab[term_name, "Pr(>|z|)"],
    or = exp(coef_tab[term_name, "Estimate"])
  )
}

run_interaction_scan <- function(
  log_mat,
  pheno_df,
  outcome_col,
  treatment_col,
  treatment_term,
  covars = NULL
) {
  mirs <- rownames(log_mat)

  map_dfr(mirs, function(mir) {
    dat <- pheno_df
    dat$mir_expr <- as.numeric(log_mat[mir, rownames(pheno_df)])

    needed_cols <- unique(c(outcome_col, treatment_col, "mir_expr", covars))
    dat <- dat |>
      dplyr::select(all_of(needed_cols)) |>
      tidyr::drop_na()

    full_rhs <- paste(c(
      paste0(treatment_col, " * mir_expr"),
      covars
    ), collapse = " + ")

    fit_all <- glm(
      as.formula(paste(outcome_col, "~", full_rhs)),
      data = dat,
      family = binomial()
    )

    treated_level <- levels(dat[[treatment_col]])[2]
    untreated_level <- levels(dat[[treatment_col]])[1]

    strat_rhs <- paste(c("mir_expr", covars), collapse = " + ")

    fit_treated <- glm(
      as.formula(paste(outcome_col, "~", strat_rhs)),
      data = dat |> filter(.data[[treatment_col]] == treated_level),
      family = binomial()
    )

    fit_untreated <- glm(
      as.formula(paste(outcome_col, "~", strat_rhs)),
      data = dat |> filter(.data[[treatment_col]] == untreated_level),
      family = binomial()
    )

    inter <- extract_glm_term(fit_all, treatment_term)
    treated <- extract_glm_term(fit_treated, "mir_expr")
    untreated <- extract_glm_term(fit_untreated, "mir_expr")

    tibble(
      miRNA = mir,
      n_all = nobs(fit_all),
      OR_interaction = inter$or,
      p_val_interaction = inter$p_value,
      se_interaction = inter$se,
      OR_treated = treated$or,
      p_val_treated = treated$p_value,
      se_treated = treated$se,
      OR_untreated = untreated$or,
      p_val_untreated = untreated$p_value,
      se_untreated = untreated$se
    )
  }) |>
    mutate(
      padj_interaction = p.adjust(p_val_interaction, method = "fdr"),
      padj_treated = p.adjust(p_val_treated, method = "fdr"),
      padj_untreated = p.adjust(p_val_untreated, method = "fdr")
    )
}

make_volcano <- function(results_df, title_text, highlight_mir) {
  plot_df <- results_df |>
    mutate(
      neglog10p = -log10(p_val_interaction),
      category = case_when(
        miRNA == highlight_mir ~ "Highlighted",
        p_val_interaction < 0.05 ~ "Significant",
        TRUE ~ "Non-significant"
      )
    )

  ggplot(plot_df, aes(x = OR_interaction, y = neglog10p)) +
    geom_point(aes(color = category, shape = category), alpha = 0.85, size = 2) +
    geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "gray50") +
    geom_text_repel(
      data = subset(plot_df, category != "Non-significant"),
      aes(label = miRNA),
      size = 3,
      max.overlaps = 30
    ) +
    scale_color_manual(values = c("Highlighted" = "blue", "Significant" = "firebrick", "Non-significant" = "gray70")) +
    scale_shape_manual(values = c("Highlighted" = 17, "Significant" = 16, "Non-significant" = 16)) +
    labs(title = title_text, x = "Odds ratio (interaction)", y = expression(-log[10](p))) +
    theme_bw(base_size = 12) +
    theme(legend.position = "top")
}

make_predprob_df <- function(log_mat, pheno_df, mir_name, outcome_col, treatment_col, treated_level) {
  dat <- pheno_df
  dat$mir_expr <- as.numeric(log_mat[mir_name, rownames(pheno_df)])
  fit <- glm(as.formula(paste(outcome_col, "~", treatment_col, "* mir_expr")), data = dat, family = binomial())

  mir_grid <- seq(quantile(dat$mir_expr, 0.1, na.rm = TRUE), quantile(dat$mir_expr, 0.9, na.rm = TRUE), length.out = 100)
  newdat <- expand.grid(mir_expr = mir_grid, tmp_group = levels(dat[[treatment_col]]))
  names(newdat)[2] <- treatment_col
  pred <- predict(fit, newdata = newdat, type = "link", se.fit = TRUE)
  newdat$fit <- pred$fit
  newdat$se <- pred$se.fit
  newdat$pred_prob <- plogis(newdat$fit)
  newdat$lower <- plogis(newdat$fit - 1.96 * newdat$se)
  newdat$upper <- plogis(newdat$fit + 1.96 * newdat$se)
  list(model = fit, pred_df = newdat)
}

plot_predprob <- function(pred_df, treatment_col, title_text) {
  ggplot(pred_df, aes(x = mir_expr, y = pred_prob, color = .data[[treatment_col]], fill = .data[[treatment_col]])) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.12, linewidth = 0) +
    geom_line(linewidth = 1) +
    labs(title = title_text, x = "miRNA expression", y = "Predicted probability", color = "Treatment", fill = "Treatment") +
    theme_bw(base_size = 12)
}

compute_auc_summary <- function(log_mat, pheno_df, mir_name, outcome_col, treatment_col, covars = NULL) {
  dat <- pheno_df
  dat$Expression <- as.numeric(log_mat[mir_name, rownames(pheno_df)])

  if (!is.null(covars)) {
    for (cv in covars) {
      if (cv %in% names(dat) && anyNA(dat[[cv]])) {
        dat[[cv]][is.na(dat[[cv]])] <- mean(dat[[cv]], na.rm = TRUE)
      }
    }
  }

  rhs_full <- paste(c(paste0(treatment_col, " * Expression"), covars), collapse = " + ")
  model_full <- glm(as.formula(paste(outcome_col, "~", rhs_full)), data = dat, family = binomial())
  pred_full <- predict(model_full, type = "response")
  roc_full <- roc(dat[[outcome_col]], pred_full, quiet = TRUE)

  stratified <- map_dfr(levels(dat[[treatment_col]]), function(g) {
    sub <- dat |> filter(.data[[treatment_col]] == g)
    rhs_sub <- paste(c("Expression", covars), collapse = " + ")
    fit <- glm(as.formula(paste(outcome_col, "~", rhs_sub)), data = sub, family = binomial())
    pred <- predict(fit, type = "response")
    roc_obj <- roc(sub[[outcome_col]], pred, quiet = TRUE)
    tibble(group = as.character(g), auc = as.numeric(auc(roc_obj)))
  })

  list(
    interaction_auc = as.numeric(auc(roc_full)),
    stratified_auc = stratified,
    model = model_full,
    roc_full = roc_full
  )
}
```

# CAMP cohort

```{r camp-load}

camp_counts <- read.csv(params$camp_counts, check.names = FALSE)
camp_pheno <- read.table(params$camp_pheno, sep = "\t", header = TRUE, check.names = FALSE)
camp_link <- read.csv(params$camp_link, check.names = FALSE)

camp_pheno <- camp_pheno[camp_pheno$TG != 2, , drop = FALSE]
camp_mat <- filter_mirnas(camp_counts, params$min_count, params$min_prop_samples)

camp_pheno_merged <- merge(camp_pheno, camp_link, by.x = "camp", by.y = "CAMP_ID", all=FALSE)
rownames(camp_pheno_merged) <- camp_pheno_merged$`row.names(CAMP_492_COUNT.Trans)`

camp_pheno_merged <- camp_pheno_merged |>
  filter(!is.na(EDHOS_cum_Y1)) |>
  mutate(
    outcome = factor(ifelse(EDHOS_cum_Y1 < 1, "low", "high")),
    outcome = relevel(outcome, ref = "low"),
    treatment = factor(TG),
    treatment = relevel(treatment, ref = "3"),
    SEX.x = factor(SEX.x)
  )

camp_pheno_merged$RACE.x = factor(camp_pheno_merged$RACE.x)
common_camp <- intersect(colnames(camp_mat), rownames(camp_pheno_merged))
camp_mat <- camp_mat[, common_camp, drop = FALSE]
camp_pheno_merged <- camp_pheno_merged[common_camp, , drop = FALSE]
```

## CAMP demographics

```{r camp-demographics}
camp_cont_vars <- c("AGE.x", "PREFEVPP_RZ", "BMI_RZ")
camp_cat_vars <- c("SEX.x", "outcome")

camp_cont_summary <- summarize_continuous_by_group(camp_pheno_merged, camp_cont_vars, "treatment")
camp_cat_summary <- summarize_categorical_by_group(camp_pheno_merged, camp_cat_vars, "treatment")

write_csv(camp_cont_summary, file.path(outdir, "demographics", "camp_continuous_summary.csv"))
write_csv(camp_cat_summary, file.path(outdir, "demographics", "camp_categorical_summary.csv"))

camp_demo_plot <- plot_demo_violin(
  camp_pheno_merged,
  vars = camp_cont_vars,
  labels = c(AGE.x = "Age", PREFEVPP_RZ = "FEV1 (% predicted)", BMI_RZ = "BMI"),
  group_col = "treatment",
  title_text = "CAMP demographics by treatment group"
)

camp_demo_plot
```

## CAMP normalization comparison

```{r camp-norm}
camp_norms <- list(
  Raw = norm_raw_log(camp_mat),
  TMM = norm_tmm_log(camp_mat),
  DESeq2 = norm_deseq2_log(camp_mat),
  Quantile = norm_quantile_log(camp_mat)
)

camp_density_df <- imap_dfr(camp_norms, ~ get_density_df(.x, .y, "CAMP"))
camp_rle_df <- imap_dfr(camp_norms, ~ get_rle_df(.x, .y, "CAMP"))
camp_meansd_df <- imap_dfr(camp_norms, ~ get_mean_sd_df(.x, .y, "CAMP"))
camp_pca_df <- imap_dfr(camp_norms, ~ run_pca_df(.x, .y, "CAMP", camp_pheno_merged, "treatment"))

p_camp_density <- plot_density(camp_density_df, "CAMP: expression distributions across normalization methods")
p_camp_rle <- plot_rle(camp_rle_df, "CAMP: RLE across normalization methods")
p_camp_meansd <- plot_mean_sd(camp_meansd_df, "CAMP: mean-SD trend across normalization methods")
p_camp_pca <- plot_pca(camp_pca_df, "CAMP: PCA across normalization methods")

p_camp_density
p_camp_rle
p_camp_meansd
p_camp_pca

ggsave(file.path(outdir, "qc", "camp_density.png"), p_camp_density, width = 12, height = 8, dpi = 300)
ggsave(file.path(outdir, "qc", "camp_rle.png"), p_camp_rle, width = 12, height = 8, dpi = 300)
ggsave(file.path(outdir, "qc", "camp_mean_sd.png"), p_camp_meansd, width = 12, height = 8, dpi = 300)
ggsave(file.path(outdir, "qc", "camp_pca.png"), p_camp_pca, width = 12, height = 8, dpi = 300)

p_camp_corr <- make_corr_plot(camp_norms$DESeq2, camp_norms$Quantile, "DESeq2 log-normalized", "Quantile log-normalized", "CAMP")
p_camp_corr

ggsave(file.path(outdir, "qc", "camp_deseq2_vs_quantile_corr.png"), p_camp_corr, width = 6, height = 5, dpi = 300)
```

## CAMP interaction analysis

```{r camp-models}
log_quantile_counts_camp <- camp_norms$Quantile

camp_results <- run_interaction_scan(
  log_mat = log_quantile_counts_camp,
  pheno_df = camp_pheno_merged,
  outcome_col = "outcome",
  treatment_col = "treatment",
  treatment_term = "treatment1:mir_expr",
  covars = c("AGE.x", "SEX.x", "RACE.x")
)

write_csv(camp_results, file.path(outdir, "models", "camp_interaction_results.csv"))

p_camp_volcano <- make_volcano(camp_results, "CAMP: miRNA × ICS interaction effects", params$highlight_mir)
p_camp_volcano

ggsave(file.path(outdir, "figures", "camp_interaction_volcano.png"), p_camp_volcano, width = 7, height = 6, dpi = 300)
```

## CAMP miR-584-5p follow-up

```{r camp-mir584}
camp_pred <- make_predprob_df(
  log_mat = log_quantile_counts_camp,
  pheno_df = camp_pheno_merged,
  mir_name = params$highlight_mir,
  outcome_col = "outcome",
  treatment_col = "treatment",
  treated_level = "1"
)

p_camp_pred <- plot_predprob(camp_pred$pred_df, "treatment", "CAMP: predicted exacerbation risk by miR-584-5p and ICS")
p_camp_pred

ggsave(file.path(outdir, "figures", "camp_mir584_predprob.png"), p_camp_pred, width = 7, height = 5, dpi = 300)

camp_auc <- compute_auc_summary(
  log_mat = log_quantile_counts_camp,
  pheno_df = camp_pheno_merged,
  mir_name = params$highlight_mir,
  outcome_col = "outcome",
  treatment_col = "treatment",
  covars = c("PREFEVPP_RZ", "AGEON", "BMI_RZ", "AGE.x", "TOT.EOS_S3", "LOG10IGE_S3")
)

write_csv(camp_auc$stratified_auc, file.path(outdir, "models", "camp_mir584_auc_stratified.csv"))
print(camp_auc$interaction_auc)
print(camp_auc$stratified_auc)
```

# CRA cohort

```{r cra-load}
cra_counts <- read.csv(params$cra_counts, check.names = FALSE)
cra_pheno <- read.csv(params$cra_pheno, check.names = FALSE)
cra_link <- read.csv(params$cra_link, check.names = FALSE)

cra_mat <- filter_mirnas(cra_counts, params$min_count, params$min_prop_samples)

cra_pheno_merged <- merge(cra_pheno, cra_link, by.x = "S_SUBJECTID", by.y = "ST.ID")
cra_pheno_merged$sample.id <- gsub("\\-", ".", cra_pheno_merged$sample.id)
rownames(cra_pheno_merged) <- cra_pheno_merged$sample.id

cra_pheno_merged <- cra_pheno_merged |>
  filter(!is.na(Hospitalized_Asthma_Last_Yr)) |>
  mutate(
    outcome = factor(Hospitalized_Asthma_Last_Yr),
    treatment = factor(Inhaled_Steroids),
    treatment = relevel(treatment, ref = "1"),
    gender.x = factor(gender.x)
  )

common_cra <- intersect(colnames(cra_mat), rownames(cra_pheno_merged))
cra_mat <- cra_mat[, common_cra, drop = FALSE]
cra_pheno_merged <- cra_pheno_merged[common_cra, , drop = FALSE]
```

## CRA demographics

```{r cra-demographics}
cra_cont_vars <- c("age.x", "pctpred_fev1_pre_BD", "bmi_pct")
cra_cat_vars <- c("gender.x", "Hospitalized_Asthma_Last_Yr")

cra_cont_summary <- summarize_continuous_by_group(cra_pheno_merged, cra_cont_vars, "treatment")
cra_cat_summary <- summarize_categorical_by_group(cra_pheno_merged, cra_cat_vars, "treatment")

write_csv(cra_cont_summary, file.path(outdir, "demographics", "cra_continuous_summary.csv"))
write_csv(cra_cat_summary, file.path(outdir, "demographics", "cra_categorical_summary.csv"))

cra_demo_plot <- plot_demo_violin(
  cra_pheno_merged,
  vars = cra_cont_vars,
  labels = c(age.x = "Age", pctpred_fev1_pre_BD = "FEV1 (% predicted)", bmi_pct = "BMI percentile"),
  group_col = "treatment",
  title_text = "CRA demographics by treatment group"
)

cra_demo_plot
```

## CRA normalization comparison

```{r cra-norm}
cra_norms <- list(
  Raw = norm_raw_log(cra_mat),
  TMM = norm_tmm_log(cra_mat),
  DESeq2 = norm_deseq2_log(cra_mat),
  Quantile = norm_quantile_log(cra_mat)
)

cra_density_df <- imap_dfr(cra_norms, ~ get_density_df(.x, .y, "CRA"))
cra_rle_df <- imap_dfr(cra_norms, ~ get_rle_df(.x, .y, "CRA"))
cra_meansd_df <- imap_dfr(cra_norms, ~ get_mean_sd_df(.x, .y, "CRA"))
cra_pca_df <- imap_dfr(cra_norms, ~ run_pca_df(.x, .y, "CRA", cra_pheno_merged, "treatment"))

p_cra_density <- plot_density(cra_density_df, "CRA: expression distributions across normalization methods")
p_cra_rle <- plot_rle(cra_rle_df, "CRA: RLE across normalization methods")
p_cra_meansd <- plot_mean_sd(cra_meansd_df, "CRA: mean-SD trend across normalization methods")
p_cra_pca <- plot_pca(cra_pca_df, "CRA: PCA across normalization methods")

p_cra_density
p_cra_rle
p_cra_meansd
p_cra_pca

ggsave(file.path(outdir, "qc", "cra_density.png"), p_cra_density, width = 12, height = 8, dpi = 300)
ggsave(file.path(outdir, "qc", "cra_rle.png"), p_cra_rle, width = 12, height = 8, dpi = 300)
ggsave(file.path(outdir, "qc", "cra_mean_sd.png"), p_cra_meansd, width = 12, height = 8, dpi = 300)
ggsave(file.path(outdir, "qc", "cra_pca.png"), p_cra_pca, width = 12, height = 8, dpi = 300)

p_cra_corr <- make_corr_plot(cra_norms$DESeq2, cra_norms$Quantile, "DESeq2 log-normalized", "Quantile log-normalized", "CRA")
p_cra_corr

ggsave(file.path(outdir, "qc", "cra_deseq2_vs_quantile_corr.png"), p_cra_corr, width = 6, height = 5, dpi = 300)
```

## CRA interaction analysis

```{r cra-models}
log_quantile_counts_cra <- cra_norms$Quantile

cra_results <- run_interaction_scan(
  log_mat = log_quantile_counts_cra,
  pheno_df = cra_pheno_merged,
  outcome_col = "outcome",
  treatment_col = "treatment",
  treatment_term = "treatment2:mir_expr",
  covars = c("age.x", "gender.x")
)

write_csv(cra_results, file.path(outdir, "models", "cra_interaction_results.csv"))

p_cra_volcano <- make_volcano(cra_results, "CRA: miRNA × ICS interaction effects", params$highlight_mir)
p_cra_volcano

ggsave(file.path(outdir, "figures", "cra_interaction_volcano.png"), p_cra_volcano, width = 7, height = 6, dpi = 300)
```

## CRA miR-584-5p follow-up

```{r cra-mir584}
cra_pred <- make_predprob_df(
  log_mat = log_quantile_counts_cra,
  pheno_df = cra_pheno_merged,
  mir_name = params$highlight_mir,
  outcome_col = "outcome",
  treatment_col = "treatment",
  treated_level = "2"
)

p_cra_pred <- plot_predprob(cra_pred$pred_df, "treatment", "CRA: predicted exacerbation risk by miR-584-5p and ICS")
p_cra_pred

ggsave(file.path(outdir, "figures", "cra_mir584_predprob.png"), p_cra_pred, width = 7, height = 5, dpi = 300)

cra_auc <- compute_auc_summary(
  log_mat = log_quantile_counts_cra,
  pheno_df = cra_pheno_merged,
  mir_name = params$highlight_mir,
  outcome_col = "outcome",
  treatment_col = "treatment",
  covars = c("pctpred_fev1_pre_BD", "`G19B_(age_onset)`", "bmi_pct", "age.x")
)

write_csv(cra_auc$stratified_auc, file.path(outdir, "models", "cra_mir584_auc_stratified.csv"))
print(cra_auc$interaction_auc)
print(cra_auc$stratified_auc)
```

# Shared signal across cohorts

```{r shared-mirs}
shared_hits <- inner_join(
  cra_results |> select(miRNA, OR_interaction_cra = OR_interaction, p_interaction_cra = p_val_interaction),
  camp_results |> select(miRNA, OR_interaction_camp = OR_interaction, p_interaction_camp = p_val_interaction),
  by = "miRNA"
) |>
  arrange(p_interaction_cra, p_interaction_camp)

write_csv(shared_hits, file.path(outdir, "models", "shared_miRNA_interaction_summary.csv"))
head(shared_hits)
```